---
title: "tissue_specific_aging"
author: "Kelaney Stalker"
date: "5/6/2022"
output: html_document
---

```{r}

#minfi code
#if using 450k data, need to change. minfi built for epic

library(minfi)
baseDir<-"CHANGE TO BASE DIRECTORY. ALWAYS KEEP QUOTATIONS."
RGset<-read.metharray.exp(base=baseDir,targets=NULL,recursive=TRUE)

Mset<-preprocessSWAN(RGset)
ratioset<-ratioConvert(Mset,what="both",keepCN=F)
z<-getBeta(ratioset)
head(z)

#next merge them

m<-read.csv("/Users/timjenkinslab/Desktop/kelaney/tissue_specific_aging/tsa_merge.csv")
head(m)
B<-t(z)
rownames(m)<-m$rn
m$rn<-NULL
head(m)
#do not do head(B) because it will be huge
k<-merge(m,B,by=0)
head(k[,c(1:10)])

#now format

rownames(k)<-k$Row.names
k$Row.names<-NULL
head(k[,c(1:10)])
K<-t(k)
head(K[,c(1:10)])

#save as RDS and then annotate 

saveRDS(K, file = "/Users/timjenkinslab/Desktop/kelaney/tissue_specific_aging/tsa_v2.rds")
B<-readRDS("/Users/timjenkinslab/Desktop/kelaney/tissue_specific_aging/tsa_v2.rds")
ann<-readRDS("/Users/timjenkinslab/Desktop/kelaney/Epic_Annotation.rds")
B<-merge(ann, B, by=0)
rownames(B)<-B$Row.names
B$Row.names<-NULL
head(B[,c(1:10)])

#save
write.csv(K, "/Users/timjenkinslab/Desktop/kelaney/tissue_specific_aging/tsa_v2.csv")

```

```{r}

#scatterplotting

j<-read.csv("/Users/timjenkinslab/Desktop/kelaney/tissue_specific_aging/jenkinsfinal.csv")
h<-read.csv("/Users/timjenkinslab/Desktop/kelaney/tissue_specific_aging/horvathfinal.csv")

plot(j$Age,j$Jenkins, col=c("blueviolet"), pch=16)
abline(lm(j$Jenkins~j$Age), lty=6)
summary(lm(j$Jenkins~j$Age))

plot(h$Age,h$Horvath, col=c("cornflowerblue"), pch=16)
abline(lm(h$Horvath~h$Age), lty=6)
summary(lm(h$Horvath~h$Age))
```

```{r}

oligosperm<-subset(j, Sample_Group== "Sperm:Oligo")
head(oligosperm)

normosperm<-subset(j, Sample_Group == "Sperm:High count")
head(normosperm)

boxplot(oligosperm$GLAD, normosperm$GLAD, xlab= "Oligo v Normo", ylab="Difference in GLAD values", main="Sperm")

oligoblood<-subset(h, Sample_Group== "Blood:Oligo")
head(oligoblood)
normoblood<-subset(h, Sample_Group == "Blood:High count")
head(normoblood)

t.test(oligoblood$GLAD, normoblood$GLAD)
```

```{r}

#boxplots 

boxplot(oligosperm$GLAD, normosperm$GLAD, col=c("lavender", "aliceblue"))
boxplot(oligoblood$GLAD.equiv, normoblood$GLAD.equiv, col=c("lavender", "aliceblue"))
boxplot(oligosperm$diff_in_pred_age, normosperm$diff_in_pred_age, col=c("lavender", "aliceblue"))
```

```{r}
# how to bar plot in ggplot 

ggplot(otwo, aes(x=Sample_Name, y=pred_age, fill=Sample_Group))+
geom_bar(stat= "identity", position = "dodge", alpha = 0.5)+
theme_bw()+
theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+ # this removes grid background
theme(axis.text.x=element_text(angle = 90, vjust = 0.5))+ # this rotates labels 90 degrees
labs(title = "Oligospermic Differences in Predicted Ages", x = "", y= "Predicted Age")

```

```{r}

#how to bar plot in base r
bo<-read.csv("~/Desktop/kelaney/tissue_specific_aging/base_oligo.csv",header=T)
head(bo)

rownames(bo)<-bo$Sample_Name
bo$Sample_Name<-NULL
bo
b0<-t(bo)

barplot(b0)
barplot(b0, beside=T)
barplot(b0, beside=T, col=rgb(.5,.8,.1))
barplot(b0, beside=T, col=c(rgb(.5,.8,.1),rgb(.1,.2,.3)))
barplot(b0, beside=T, col=c(rgb(.5,.8,.1),rgb(.1,.2,.3)))
legend("topright",c("Blood","Sperm"),pch=c(15,15),col=c(rgb(.5,.8,.1),rgb(.1,.2,.3)),bty="n")

barplot(b0, beside=T, col=c(rgb(.5,.8,.1),rgb(.1,.2,.3)),ylim=c(0,60))
legend("topright",c("Blood","Sperm"),pch=c(15,15),col=c(rgb(.5,.8,.1),rgb(.1,.2,.3)),bty="n")




